---
Title: "A3: Comparisons"
Author: "Harshali Singh, Vishal Mehta"
Output: 
  pdf_document:
    toc: true
    number_sections: true
---

### Link Referred for Fast median 

https://en.wikipedia.org/wiki/Quickselect

###Configurations

We have considered 3 Statistics for 5 versions of code, for a total of 15 configurations:-

(i) singe threaded Java
(ii) multi-threaded Java
(iii) pseudo-distributed MR
(iv) distributed MR
(v) Scala

Statistics:

(i) Mean
(ii) Median
(iii) Fast Median

###Implementation

We have calculated different statistics in the following way:-

####Mean
The Mapper code reads the input files as <Key,Value> pairs and emits key value pairs. We are cleaning each file and performing sanity check in the Mapper. The Key is carrier code and month separated by tab.

The reducer code then sums all the average ticket prices for each carrier code and take the average using the total count of that particular key.

####Median
The Mapper code reads the input files as <Key,Value> pairs and emits key value pairs. We are cleaning each file and performing sanity check in the Mapper. The Key is carrier code and month separated by tab.

The reducer code then sorts all the average ticket prices of a particular key and finds the middle value as the median price.


####Fast Median
The Mapper code reads the input files as <Key,Value> pairs and emits key value pairs. We are cleaning each file and performing sanity check in the Mapper. The Key is carrier code and month separated by tab.

The reducer code then finds the median price by using QuickSelect algorithm which has a running time of O(n). 

Assumptions: We have only considered Average Ticket Prices for 2015 flights.


###Benchmarking Harness

We calculate execution time of each configuration in sequence and dump the time in a csv file. We achieve this using the Linux Makefile script where each configuration is run in sequence and time is fetched. We are using R to read the csv files and plot a graph.

```{r, echo=FALSE}

library(ggplot2)

#Read the output file
table = read.csv("benchmarking.csv", sep=",")

```

###Graph Plot

We are using R to plot the executing timings of all 15 configurations. 

```{r}

#Plot configuration vs timing and for each statistic calculated
p <- ggplot(table, aes(CONF, TIME)) + geom_point()
p + facet_grid(. ~ MODE)


```

### Sample Output:

This is the output of MeanComparison with a sample input of one file.

1	AA	43971	530.1824
1	AS	13213	230.10681
1	B6	21579	523.64746
1	DL	64082	598.80695
1	EV	49589	305.67844
1	F9	6824	131.74747
1	HA	6434	302.6006
1	MQ	29635	288.26422
1	NK	8730	487.5079
1	OO	47737	307.34674
1	UA	38324	969.1005
1	US	33424	600.1598
1	VX	4726	652.921
1	WN	99553	159.35498

###Conclusion
The output of this assignment is a Report.pdf file which explains our implementation and the plot. Also it includes the sample output with a sample input of the MeanComparison Map reduce program.
The Fast median version runs faster than the normal median with the same input and same output values.  
Moreover, the single threaded version runs slower than the Multi threaded version with a given input.

